---
title: "Odorant Clustering"
output: html_notebook
---

```{r, loading}
# tidyverse
require(webchem)
require(factoextra)
require(ggrepel)
require(stringr)
require(broom)
require(forcats)
require(magrittr)
require(tidyverse)

theme_pub <- function(base_size = 13, base_family = "") {
  require(grid)
  # based on https://github.com/noamross/noamtools/blob/master/R/theme_nr.R
  # start with theme_bw and modify from there!
  theme_bw(base_size = base_size, base_family = base_family) +# %+replace%
    theme(
      # grid lines
      panel.grid.major.x = element_line(colour="#ECECEC", size=0.5, linetype=1),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_line(colour="#ECECEC", size=0.5, linetype=1),
      panel.background   = element_blank(),
      
      # axis options
      axis.ticks.y   = element_blank(),
      axis.title.x   = element_text(size=rel(2.25), vjust=0.25),
      axis.title.y   = element_text(size=rel(2.25), vjust=0.35),
      axis.text      = element_text(color="black", size=rel(1.5)),
      
      # legend options
      legend.title    = element_blank(),
      legend.key      = element_rect(fill="white"),
      legend.key.size = unit(1, "cm"),
      legend.text     = element_text(size=rel(2)),
      
      # facet options
      strip.text = element_text(size=rel(2)),
      
      # title options
      plot.title = element_text(size=rel(3), vjust=0.25, hjust=0.5)
    )
}

# set the theme and brewer color
theme_set(theme_pub())
```

# Autoencoder Prep

Load the finallized list of chemicals and their properties (InChIKey, SMILES, etc) and join with our hits

```{r, or-hits}
# Hand cleaned data-set
final.chems <- read_tsv('./final-chems.tsv')

# Rishi's data (fix some names to key)
or.hits <- read_csv('AllORHits.csv') %>%
  mutate(Odorant = case_when(
    Odorant == 'ethyl maltol' ~ 'Ethyl Maltol',
    Odorant ==  " (R)-(+)-Limonene" ~ '(R)-(+)-Limonene', # still doesnt quite work..
    Odorant == 'Decanoic_Acid' ~ 'Decanoic Acid',
    Odorant == 'Bis(2-methyl-3-furyl)disulphide)' ~ 'Bis(2-methyl-3-furyl)disulphide',
    Odorant == '2_coumaranone' ~ '2-Coumaranone',
    TRUE ~ Odorant
  ))

# output the or's excluding the NA's
or.hits %>% 
  rename(name = Odorant) %>% 
  left_join(final.chems) %>% 
  select(OR, name, min_activating_conc, InChIKey, InChI) %>%
  filter(!is.na(InChIKey)) %>% 
  arrange(OR) %>%
  write_csv('./ors-inchi-rishi.csv')
```

We then output this to a nice CSV that our python script can read in to plot the structures and their minimum activating concentration.

# Global Chemical Space Analysis
Let's see where our chemicals map globally. Make sure to run everything through the autoencoder (`./autoencoder_ipython/autoencoder.py`) first

```{r, global-landscape}
global.pca <- read_csv('./latent-space-pca.csv.gz') %>%
  rename(name = X1) %>%
  mutate(Lib = if_else(str_detect(name, 'CHEM'), 'No', 'Yes')) 

# where do our chemicals lay in space
global.pca %>%
  mutate(Lib = if_else(str_detect(name, 'CHEM'), 'No', 'Yes')) %>% 
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(alpha=0.05, color = "#64acbe") +
  ggrepel::geom_text_repel(
    data=global.pca %>% filter(Lib == 'Yes'), 
    aes(label = name), 
    force=3,
    segment.color = '#e4acac',
    max.iter=5000,
    box.padding = 0.5
  ) +
  geom_point(data=. %>% filter(Lib == 'Yes'), color='red') +
  labs(
    x = 'Chemical Space PC1',
    y = 'Chemical Space PC2'
  )

# where do our hits lay wrt the global space for each or?
# same scheme as heatmap
or.hits %>%
  rename(name = Odorant, act = min_activating_conc) %>% 
  inner_join(global.pca, by = 'name') %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(
    data=global.pca %>% filter(Lib == 'Yes'), 
    pch=21, 
    fill='#EFF3FF',
    color='black'
  ) +
  geom_point(
    pch=21, 
    aes(fill=as.character(act))
  ) +
  facet_wrap(~OR) +
  scale_fill_manual(values = c('#08306b','#4292c6','#bdd7e7')) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.background = element_blank(),
    legend.position = 'bottom'
  )

# nates color
or.hits %>%
  rename(name = Odorant, act = min_activating_conc) %>% 
  inner_join(global.pca, by = 'name') %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(data=global.pca %>% filter(Lib == 'Yes'), color='grey') +
  geom_point() + #aes(color=as.character(act))) +
  facet_wrap(~OR) +
  # scale_color_manual(values = c('#225ea8','#41b6c4','#a1dab4')) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.background = element_blank(),
    legend.position = 'bottom'
  )
```

# Appendix
## Converting Names
First we will convert the names into unique identifiers using a variety of chemical lookup services. We will use this as a jumping off and manually edit the rest.

```{r, name-convert}
or.chems <- read_csv('ChemicalList.csv')

# search pubchem by names
pubchem <- or.chems %$% 
  get_cid(`Chemical Name`) %>%
  unlist() %>% 
  enframe() %>% 
  nest() %>% 
  mutate(foo = map(data, ~pc_prop(.x$value, properties = c('InChIKey', 'InChI', 'CanonicalSMILES', 'IsomericSMILES')))) %>%
  unnest() %>%
  select(-value)

write_csv(pubchem, '~/Downloads/pubchem-eric.csv')

# use the chemical translation service
CTS <-  or.chems %$% 
  cts_convert(`Chemical Name`, 'Chemical Name', 'InChIKey') %>%
  unlist() %>% 
  enframe()

# compare the two
pubchem %>% 
  full_join(CTS) %>% 
  rename(InChIKey_CTS = value) %>%
  mutate(Diff = if_else(InChIKey == InChIKey_CTS, 'Same', 'Diff')) %>% 
  select(name, PubChemID=CID, InChIKey, InChIKey_CTS, 
         Diff, InChI, IsomericSMILES, CanonicalSMILES) %>%
  write_csv('~/Downloads/eric-cts-vs-pubchem.csv')
```

## Hit Space Analysis


```{r}
latent.mat <- read_csv('./chem-vecs.csv', col_names = T)

# hclust
latent.mat %>% 
  as.data.frame() %>%
  column_to_rownames('X1') %>%
  dist() %>%
  hclust() %>%
  factoextra::fviz_dend()

# PCA
latent.mat %>% 
  as.data.frame() %>%
  column_to_rownames('X1') %>%
  prcomp %>%
  factoextra::fviz_pca_ind(repel = T)

# get the two principle components
chem.coords <- latent.mat %>% 
  as.data.frame() %>%
  column_to_rownames('X1') %>%
  prcomp %$%
  x[,1:2] %>%
  as.data.frame() %>%
  rownames_to_column(var='name') %>%
  as_tibble() 

or.hits %>%
    rename(name = Odorant) %>%
    inner_join(chem.coords) %>%
    ggplot(aes(x=PC1, y=PC2)) +
    geom_point(data=chem.coords, color='grey') +
    geom_point() +
    facet_wrap(~OR)
```

We can also try the dreaded t-sne

```{r, TSNE}
SNE <- latent.mat %>% 
  as.data.frame() %>%
  column_to_rownames('X1') %>%
  Rtsne::Rtsne() %$%
  Y %>%
  as_tibble() %>%
  bind_cols(latent.mat %>% select(name = X1))

or.hits %>%
    rename(name = Odorant) %>%
    inner_join(SNE) %>%
    ggplot(aes(x=V1, y=V2)) +
    geom_point(data=SNE, color='grey') +
    geom_point() +
    facet_wrap(~OR)
```
